Nodal Discontinuous Galerkin Method for the Two-Moment Model of Neutrino Transport

### 2017-06-02 Documents/M1/dgM1Curvilinear.tex
Modification backup for chapter 6.

%\ee{Ran write about tabulated B85 absorption and scattering opacities.}

Given by \citet{bruenn_1985}, the absorption opacity of neutrino $\chi$ 
\begin{equation}
  \tilde{\chi}(\epsilonNu,\bU) = \chi(\epsilonNu,\bU) + \eta(\epsilonNu,\bU), 
\end{equation}
where $\eta(\epsilonNu,\bU) $ is the neutrino emissivity
\begin{equation}
\eta = \eta_{nucleon} + \eta_{nuclear},
\end{equation}  
and $ \chi(\epsilonNu,\bU)$ is the neutrino absorptivity, which is the inverse of neutrino absorption mean free path.
They have the following relation:
\begin{equation}
\chi(\epsilonNu,\bU) = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T} \eta(\epsilonNu,\bU). 
\end{equation}

Therefore,
\begin{equation}
\chi = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T} (\eta_{nucleon} + \eta_{nuclear} ) = \chi_{nucleon} + \chi_{nuclear}.
\end{equation}
The two terms of electron-neutrino absorption on nucleons are
 \begin{align}
    \chi_{nucleon}(\epsilonNu,\bU) & = \dfrac{G^2}{\pi} \eta_{np}(g_V^2+3g_A^2)[1-F_e(\varepsilon+Q)](\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonabsorption} ,\\
    \eta_{nucleon}(\epsilonNu,\bU) &= \dfrac{G^2}{\pi}\eta_{pn}(g_V^2+3g_A^2)F_e(\varepsilon+Q)(\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonemissivity} ,
\end{align}
and the two terms of electron-neutrino absorption on nuclear are
\begin{align}
    \chi_{nuclear}(\epsilonNu,\bU) & = \dfrac{G^2}{\pi} n_A e^{(\mu_n -\mu_p -Q')/k_B T} g^2_A \dfrac{2}{7} N_p(Z) N_h(N)[1-F_e(\varepsilon+Q')](\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearabsorption} ,\\
    \eta_{nuclear}(\epsilonNu,\bU) &=\dfrac{G^2}{\pi} n_A g^2_A \dfrac{2}{7} N_p(Z) N_h(N)F_e(\varepsilon+Q')(\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearemissivity},
\end{align} 
where $g_V = 1$ and $g_A =1.23$ are form factors resulting from virtual strong-interaction processes, $M_e = 0.511$ MeV is the electron rest energy, $Q = 1.2935$ MeV is the difference between the neutron and proton rest energies, $\mu_n$ and $\mu_p$ are the neutron and proton chemical potential, respectively. And $Q'$ is defined by $Q' = \mu_n -\mu_p + \Delta$, where $\Delta=3$MeV is the energy of the $1f_{5/2}$ level above the ground state.

@@

The sum of Eq.~\eqref{eq:nucleonabsorption}, ~\eqref{eq:nucleonemissivity}, ~\eqref{eq:nuclearabsorption} and ~\eqref{eq:nuclearemissivity} gives the explicit expression of the absorption opacity of neutrino $\chi$ .

The total opacity $\kappa(\epsilonNu,\bU)=\chi(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$, and $\sigma(\epsilonNu,\bU)$ is the scattering opacities. The scattering opacities $\sigma(\epsilonNu,\bU)$ contains isoenergetic scattering opacities and have a integral form in the Boltzmann equation: if we include emission, absorption, and isoenergetic scattering of neutrino by nucleons and nuclei, the simplified Boltzmann equation is
\begin{align}
\dfrac{\partial f(\epsilonNu,\bU)}{\partial t} = \eta(\epsilonNu,\bU) - \tilde{\chi}(\epsilonNu,\bU)f(\epsilonNu,\bU) & + \dfrac{1}{h^3} \epsilonNu^2 \int d\mu^\prime_0 R_{IS}( \mu^\prime_0, \mu_0,\epsilonNu, \bU) f(\mu^\prime_0, \epsilonNu,\bU) \nonumber \\
	& - \dfrac{1}{h^3} \epsilonNu^2 f(\epsilonNu,\bU)\int d\mu^\prime_0 R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) .
\end{align}

The last two terms are contributions from isoenergetic scattering of neutrino by nucleons and nuclei. The isoenergetic scattering kernel $R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)$ has the following relation:
\begin{align}
R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) & = R_{IS}^{nucleons}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) +  R_{IS}^{nuclei}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)\\
& = \int_{0}^{2\pi}d
\end{align}

>>>>

%\ee{Ran write about tabulated B85 absorption and scattering opacities.}
In this section, we bring the speed of light $c$ back.

The total opacity $\kappa(\epsilonNu,\bU)=\tilde{\chi}(\epsilonNu,\bU)+\sigma(\epsilonNu,\bU)$, where $\tilde{\chi}(\epsilonNu,\bU)$ is the absorption opacity of neutrino and  $\sigma(\epsilonNu,\bU)$ is the scattering opacities. 
The absorption opacity of neutrino $\tilde{\chi}(\epsilonNu,\bU)$ 
\begin{equation}
  \tilde{\chi}(\epsilonNu,\bU) = \chi(\epsilonNu,\bU) + \eta(\epsilonNu,\bU), 
\end{equation}
 is the sum of the neutrino emissivity $\eta(\epsilonNu,\bU) $ and the neutrino absorptivity  $ \chi(\epsilonNu,\bU)$. The scattering opacities $\sigma(\epsilonNu,\bU)$ contains isoenergetic scattering opacities, neutrino-electron scattering opacities and have a integral form in the Boltzmann equation: if we include emission, absorption, isoenergetic scattering of neutrino by nucleons and nuclei, and neutrino electron scattering, the simplified Boltzmann equation is
\begin{align}
\dfrac{1}{c}\dfrac{\partial f(\epsilonNu,\bU)}{\partial t} = 
&  \eta(\epsilonNu,\bU) - \tilde{\chi}(\epsilonNu,\bU)f(\epsilonNu,\bU) \nonumber \\
& + \dfrac{1}{c(hc)^3} \epsilonNu^2 \int d\mu^\prime_0 R_{IS}( \mu^\prime_0, \mu_0,\epsilonNu, \bU) f(\mu^\prime_0, \epsilonNu,\bU) \nonumber \\
& - \dfrac{1}{c(hc)^3} \epsilonNu^2 f(\epsilonNu,\bU)\int d\mu^\prime_0 R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) \nonumber \\
& + \dfrac{1}{c(hc)^3} [1- f(\epsilonNu,\bU)] \int d\epsilonNu^{\prime} \epsilonNu^{\prime 2} \int d\mu^\prime_0 R^{in}_{NES} (\mu_0,\mu^\prime_0,\epsilonNu,\epsilonNu^{\prime} )f(\mu^\prime_0,\epsilonNu^{\prime} ) \nonumber \\
& - \dfrac{1}{c(hc)^3} f(\epsilonNu,\bU) \int d\epsilonNu^{\prime} \epsilonNu^{\prime 2} \int d\mu^\prime R^{out}_{NES} (\mu_0,\mu^\prime_0,\epsilonNu,\epsilonNu^{\prime} ) [1-f(\mu^\prime_0,\epsilonNu^{\prime} )] .
\end{align}

The third and fourth terms are contributions from isoenergetic scattering of neutrino by nucleons and nuclei. The last two terms are contributions from neutrino-electron scattering.

\subsection{The neutrino absorptivity}

Given by \citet{bruenn_1985}, the absorption opacity of neutrino $\tilde{\chi}(\epsilonNu,\bU)$ 
\begin{equation}
\tilde{\chi}(\epsilonNu,\bU) = \chi(\epsilonNu,\bU) + \eta(\epsilonNu,\bU), 
\end{equation}
where $\eta(\epsilonNu,\bU) $ is the neutrino emissivity
\begin{equation}
\eta = \eta_{nucleon} + \eta_{nuclear},
\end{equation}  
and $ \chi(\epsilonNu,\bU)$ is the neutrino absorptivity, which is the inverse of neutrino absorption mean free path.
They have the following relation:
\begin{equation}
\chi(\epsilonNu,\bU) = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T} \eta(\epsilonNu,\bU). 
\end{equation}

Therefore,
\begin{equation}
\chi = e^{(\epsilonNu-\mu_{\nu})/k_{\mbox{\tiny B}}T} (\eta_{nucleon} + \eta_{nuclear} ) = \chi_{nucleon} + \chi_{nuclear}.
\end{equation}

The two terms of electron-neutrino absorption on nucleons are
\begin{align}
\chi_{nucleon}(\epsilonNu,\bU) & = \dfrac{G^2}{\pi} \eta_{np}(g_V^2+3g_A^2)[1-F_e(\varepsilon+Q)](\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonabsorption} ,\\
\eta_{nucleon}(\epsilonNu,\bU) &= \dfrac{G^2}{\pi}\eta_{pn}(g_V^2+3g_A^2)F_e(\varepsilon+Q)(\varepsilon+Q)^2[1-\dfrac{M^2_e}{(\varepsilon+Q)^2}]^{1/2} \label{eq:nucleonemissivity} ,
\end{align}
and the two terms of electron-neutrino absorption on nuclear are
\begin{align}
\chi_{nuclear}(\epsilonNu,\bU) & = \dfrac{G^2}{\pi} n_A e^{(\mu_n -\mu_p -Q')/k_B T} g^2_A \dfrac{2}{7} N_p(Z) N_h(N)[1-F_e(\varepsilon+Q')](\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearabsorption} ,\\
\eta_{nuclear}(\epsilonNu,\bU) &=\dfrac{G^2}{\pi} n_A g^2_A \dfrac{2}{7} N_p(Z) N_h(N)F_e(\varepsilon+Q')(\varepsilon+Q')^2[1-\dfrac{M_e^2}{(\varepsilon+Q')^2}]^{1/2} \label{eq:nuclearemissivity},
\end{align} 
where $g_V = 1$ and $g_A =1.23$ are form factors resulting from virtual strong-interaction processes, $M_e = 0.511$ MeV is the electron rest energy, $Q = 1.2935$ MeV is the difference between the neutron and proton rest energies, $\mu_n$ and $\mu_p$ are the neutron and proton chemical potential, respectively. And $Q'$ is defined by $Q' = \mu_n -\mu_p + \Delta$, where $\Delta=3$MeV is the energy of the $1f_{5/2}$ level above the ground state.

@@

The sum of Eq.~\eqref{eq:nucleonabsorption}, ~\eqref{eq:nucleonemissivity}, ~\eqref{eq:nuclearabsorption} and ~\eqref{eq:nuclearemissivity} gives the explicit expression of the absorption opacity of neutrino $\chi(\epsilonNu,\bU)$ .

\subsection{The isoenergetic scattering kernel}
The isoenergetic scattering kernel $R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU)$ has the following relation:
\begin{align}
R_{IS}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) & = R_{IS}^{nucleons}(\mu_0, \mu^\prime_0, \epsilonNu,\bU) +  R_{IS}^{nuclei}(\mu_0, \mu^\prime_0, \epsilonNu,\bU),
\end{align}
where
\begin{align}
\cos\theta = \mu_0\mu^\prime_0 + [(1-\mu_0^2)(1-\mu_0^{\prime 2})]^{1/2}\cos\theta,
\end{align}
and
\begin{align}
R_{IS}(\cos\theta, \epsilonNu, \bU) = R^{nucleons}(\cos\theta,\epsilonNu, \bU) + R^{nuclei}_{IS}(\cos\theta, \epsilonNu, \bU)
\end{align}
where
\begin{align}
& R^{nucleons}_{IS} (\cos\theta,\epsilonNu, \bU) = \dfrac{(2\pi)^2 G_F^2}{h} \nonumber \\
& \cdot \left\lbrace  \eta_{nn}\left[ (h^n_V)^2 + 3(h^n_A)^2\right] + \eta_{pp} \left[ (h^p_V)^2 + 3(h^p_A)^2\right] +\eta_{nn}\left[ (h^n_V)^2 - (h^n_A)^2\right]\cos\theta + \eta_{pp}\left[ (h^p_V)^2 - (h^p_A)^2\right]\cos\theta \right\rbrace 
\end{align}
and
\begin{align}
R^{nuclei}_{IS}(\cos\theta,\epsilonNu,\bU)  = \dfrac{(2\pi)^2 G_F^2}{h}\dfrac{\rho X_H}{m_B}A\left(  C_{V0} - \dfrac{1}{2}\dfrac{N-Z}{A} C_{V1}\right)^2(1+\cos\theta)e^{-4b\epsilonNu^2(1-\cos\theta)}.
\end{align}
$h^p_V$, $h^n_V$, $h^p_A$ and $h^n_A$, are the nucleon neutral current form factors arising from virtual strong-interaction processes: $h_V^p = \dfrac{1}{2} - 2 \sin^2\theta_W$, $h^n_V = - \dfrac{1}{2}$, $h_A^p = \dfrac{1}{2}g_A$, and $h^n_A = - \dfrac{1}{2} g_A$, where $\theta_W$ is the Weinberg angle. The constants, $C_{V0}$ and $C_{V1}$, are defined in terms of $h_V^p$ and $h_V^n$: $C_{V0} = \dfrac{1}{2}(h_V^p + h_V^n)$ and $C_{V1} = h_V^p - h_V^n$. The quantity $b$ is given by $b = 4.8 \times 10^{-6} A^{2/3}$.

For nondegenerate, nonrelativistic nucleons, $\eta_{pp}$ and $\eta_{nn}$ are given by
\begin{align}
&\eta_{pp} = n_p \\
&\eta_{nn} = n_n.
\end{align}
